Towards a self healing swarm through the removal of voids and anomalies

### 2018-07-16 paper.tex
Draft 1

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\normalsize,
  numbers=left,
  numberstyle=\normalsize,
  stepnumber=1,
  xleftmargin=2em
}
\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\let\latexvec=\vec
\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Towards a self healing swarm through the removal of voids and anomalies}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, Michael Brockway\authorrefmark{1}, Alun Moon\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}

\markboth
{Eliot \headeretal: Towards a self healing swarm through the removal of voids and anomalies}
{Eliot \headeretal: Towards a self healing swarm through the removal of voids and anomalies}

@@

\begin{abstract}
%\boldmath
Swarms consist of many agents that interact using a simple set of rules that create emergent behaviours. One such behaviour makes a swarm tolerant of failures. Failures occur for may reasons; for example: resource exhaustion, component failure, or a disruption from an external event. The loss of agents reduces the size of a swarm and changes the structure in places where the agents are lost. A swarm's structure can also be irregular from an initial deployment or due to an obstacle in its path. These changes in the structure and/or size do not stop a swarm from functioning but can affect the outcome of a task to which the swarm is being applied. The paper introduces and demonstrates an emergent behaviour that can be induced into a swarm that can counter the effect of agent loss or irregular structures within a swarm. The behaviour causes concave deformities to be removed creating a healing-effect. This technique is implemented by each agent identifying it's position in relation to it's neighbours. If the agent meets a specific criteria the agent's `normal' movement is altered in such a way as to remove a gap between two of it's neighbours. These changes in direction of agents that meet the criterion cause the swarm, as a whole, to be reorganised into a more uniform structure with no internal voids.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{IEEEkeywords}
Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
\end{IEEEkeywords}

\titlepgskip=-15pt
\maketitle

\section{Introduction}\label{sec:ConcaveReduction}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively.
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two \textit{field effects} to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \textit{Field effects} have ranges around an agent to determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects})the effects are derived from these ranges as vectors. All ranges must fall within the sensing capabilities of the agent ($S_b$). It is usual for the cohesion field to have a radius ($C_b$) which is larger than the repulsion radius ($R_b$). When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will tend towards moving away from the obstacle.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/SwarmStableShape1}
{Stable swarm edges\label{concave:SwarmStableShape1}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=5cm]{figures/SwarmStableShape1}
%% \end{center}
%% \caption{Stable swarm edges\label{concave:SwarmStableShape1}}
%% \end{figure}

@@

\section{Related Work}\label{sec:RelatedWork}
Research have identified several approaches to resolving swarm structure issues. A prototype framework for self healing swarms was developed by Dai et al. which considered the problem of agent failure in hostile environments~\cite{DHMRZ:06}. This is in line with work carried out by Vassev and Hinchey who model swarm deployment using ASSL (Autonomic System Specification Language)~\cite{VH:09}. This technique was used by NASA (National Aeronautics and Space Administration) when developing their ANTS (Autonomous Nano Technology Swarm) for use in asteroid belt exploration. This line of research is more focused towards an agent's internal systems failure rather than the removal of anomalies in a swarm distribution. Roach et al. takes a different approach and focuses on the effects of sensor failure and the effect that has on agent distribution~\cite{RMT:15}. The closest research to concave reduction, as discussed in this paper, is that of Lee and Chong who identify the issue of concave edges within swarms in an attempt to create regular lattice formations~\cite{GN:08}. The main focus of their paper is the restructuring of the internal distribution of inter-agent formations. The research by Ismail and Timmis used the approach of \textit{bio-inspired} healing using \textit{granuloma formation}, a biological method for encapsulating an antigen~\cite{IT:10}.

Concave reduction is an extension of the work discussed by Ismail and Timmis in their paper on `self-healing' \cite{IT:10} and Lee and Chong's work on identifying concave edges~\cite{GN:08}. The technique also draws on the work of McLurkin and Demaine who have developed algorithms to detect perimeter types~\cite{MD:09}, although this paper does not identify the perimeter types as this would require a communications infrastructure.

\section{Perimeter Detection}\label{sec:PerimeterDetection}
%In a 2D Euclidean plane a circle provides the maximum area with the minimum perimeter size. A circles perimeter is convex in nature.

In a disorganised swarm agents that are on a perimeter edge need to be identified as the starting point of coordinating the structure. The edges can form part of an outer (green) or inner edge (red) (Fig.~\ref{fig:PerimeterBots1}) or be part of an island~(Fig.~\ref{fig:PerimeterBots2}).

Agents are identified as being perimeter-based if they comply with a set of conditions. These conditions are: 
\begin{itemize}
  \item There are at least two adjacent neighbours that cannot detect each other.
  \item The agent has less than 4 neighbours.
  \item The angle between two neighbours is $> 180^\circ$.
\end{itemize} 
These conditions create an identifiable set of agents that create a perimeter.

This process detects both the outer edge of a swarm and any internal features~(voids) that satisfy the same set of conditions. It is therefore possible to have both voids and islands of agents within the same swarm~(Fig.~\ref{fig:PerimeterBots2}). Voids are best defined as perimeters that are concave in nature and inside another perimeter. McLurkin describes two types of perimeters convex and concave~\cite{MD:09}. A convex perimeter is an edge where the average angle of the agent's exposed faces are~$< 180^\circ$. A concave perimeter is where the average exposed angle is~$> 180^\circ$.

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/PerimeterBots2}
{Inner Island Perimeter\label{fig:PerimeterBots2}}

%\subsection{Full-perimeter detection}\label{sec:PerimeterAgentDetection} 
The process of detecting perimeter agents involves several steps. Primarily it involves identifying agents that are not surrounded by interconnected neighbours, as shown in Fig.~\ref{reduced:Perimeter1}, where $b^0$ and $b^1$ are outside of each others cohesion range and are neighbours of $b$. These processes are discussed in more detail here.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/Perimeter1}
{Agent with detected neighbours\label{reduced:Perimeter1}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/Perimeter1}
%% \end{center}
%% \caption{Complex multifaceted agents \label{reduced:Perimeter1}}
%% \end{figure}

The first check is a short circuit count of the number of neighbours. If an agent has $\leq 4$ neighbours and the swarm parameters are for a hexagonal swarm configuration then the agent is a perimeter agent. This identification process is shown in~Fig.~\ref{algo:SmallNeighbourCount}.

@@

If the short circuit \textit{neighbour count} returns false then further conditions must be checked. The additional checks require a dictionary of agent/angular values to be generated as shown in Fig.~\ref{eq:ConcaveVoidPhysics1}. The algorithm shown in Fig.~\ref{algo:getNeighbourAngles} shows how the dictionary of agent/angle values are produced.

% \begin{equation}
% S_b \buildrel \Delta \over = \{(b',\angle\{(b'~b~b^0)) : b' \in nbr(b)\}\}
% \label{eq:ConcaveVoidPhysics1}
% \end{equation}

Equation~\ref{eq:ConcaveVoidPhysics1} generates the dictionary set of all the neighbours along with an angle that each of the agent's neighbour's make with the first detected neighbour.
%% 
%% \begin{center}
%% \begin{equation}‎
%% \label{eq:ConcaveVoidPhysics2}
%% G_b \buildrel \Delta \over =‎ gap(S_b,\angle<180^\circ)
%% \end{equation}‎
%% \end{center}

@@

Using the angular dictionary from Fig.~\ref{algo:getNeighbourAngles} it is possible to identify if adjacent neighbours can detect each other. If the agent is within all of its neighbours boundary then it is not on the perimeter~(Fig.~\ref{algo:CheckVisibility}).

As all the agents have the same field effects the `visibility' of two neighbours can be determined. The angle and distance of each neighbour pair allows the cosine rule to be used to calculate the distance the neighbours are apart as shown in~\ref{eq:CosineRule1} and Fig.~\ref{reduced:Triangulate}. The distance is checked against the neighbour range, if the distance is $\leq N_b$ where $N_b$ is the \textit{neighbour field} then it can be assumed that the neighbours have `sight' of each other.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=3cm]{figures/Triangulate}
{Neighbour visibility\label{reduced:Triangulate}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=3cm]{figures/Triangulate}
%% \end{center}
%% \caption{Neighbour visibility\label{reduced:Triangulate}}
%% \end{figure}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/Perimeter2}
{Convex agents\label{reduced:Perimeter2}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/Perimeter2}
%% \end{center}
%% \caption{Convex multifaceted agents \label{reduced:Perimeter2}}
%% \end{figure}

@@

\section{Concave Reduction Application}\label{sec:ConcaveReductionApplication}
In a static swarm, where there are no \textit{destination vectors}, concave reduction will result in a restructuring motion that creates a more `rounded' swarm~(Fig.~\ref{fig:OuterPerimeter3}). Concave reduction also creates a surround effect as it removes voids from a swarm~(Section~\ref{voids:ObjectSurrounding}). When concave reduction is applied to a mobile swarm the void closing effect can be used to improve reconnaissance coverage as a swarm passes around obstacles~(Section~\ref{concave:mobileSwarm1}).

Although these effects improve the potential applications of swarms there is a negative impact introduced by concave reduction. In some circumstances concave reduction can create an artificial aggregate \textit{destination vector}.

To implement void reduction full perimeter detection is required. Concave reduction does not require the perimeter type to be identified and therefore no communications infrastructure is required~\cite{MD:09,NIM:09,SOM:12,ZFG:13,JG:13}. The main limitation a communications infrastructure imposes on a swarm is the number of agents a swarm can contain. This is due to the propagation time of agent positions throughout the swarm.

The identification of candidate agents on the perimeter is incorporated into the perimeter detection algorithm as shown in Fig.~\ref{algo:checkVisibility2}.\\

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/PerimeterBots4}
{Link\label{fig:Connector1}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/PerimeterBots4}
%% \end{center}
%% \caption{Link\label{fig:Connector1}}
%% \end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/PerimeterBots5}
{Bridge\label{fig:Connector2}}

%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/PerimeterBots5}
%% \end{center}
%% \caption{Bridge\label{fig:Connector2}}
%% \end{figure}

\section{Concave reduction agent movement}\label{concave:AgentMovement}
Adding a further characteristic to the motion of a swarm necessitates a revision to the existing agent model~(Equation.~\ref{eq:InterAgentMovement2}). With concave reduction this revision is based upon the identification of the concave perimeter edges as shown in~Fig.~\ref{concave:VoidConcave1} as $(b^{'},b,b^{''})$.

Equation~\ref{eq:InterAgentMovement2} identifies the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.

@@

Where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:BaselineConcaveReduction}.\\

% \begin{figure}
% \begin{center}
% \includegraphics[width=5cm]{figures/VoidConcave1}
% \end{center}
% \caption{Agent concave motion\label{concave:VoidConcave1}}
% \end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/VoidConcave1}
{Agent concave motion\label{concave:VoidConcave1}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/InterAgentEffect1}
{Initial Position\label{fig:InterAgentEffect1}}

>>>>

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\normalsize,
  numbers=left,
  numberstyle=\normalsize,
  stepnumber=1,
  xleftmargin=2em
}
\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\let\latexvec=\vec
\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
\newcommand{\card}[1]{\vert{#1}\vert}
\newcommand{\magn}[1]{\Vert{#1}\Vert}
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Towards a self healing swarm through the removal of voids and anomalies}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, Michael Brockway\authorrefmark{1}, Alun Moon\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}
\markboth
{Eliot \headeretal: Towards a self healing swarm through the removal of voids and anomalies}
{Eliot \headeretal: Towards a self healing swarm through the removal of voids and anomalies}

@@

\begin{abstract}
%\boldmath
Swarms consist of many agents that interact using a simple set of rules that create emergent behaviours. One such behaviour makes a swarm tolerant of failures. Failures occur for may reasons; for example: resource exhaustion, component failure, or a disruption from an external event. The loss of agents reduces the size of a swarm and changes the structure in places where the agents are lost. A swarm's structure can also be irregular from an initial deployment or due to an obstacle in its path. These changes in the structure and/or size do not stop a swarm from functioning but can affect the outcome of a task to which the swarm is being applied. The paper introduces and demonstrates an emergent behaviour that can be induced into a swarm that can counter the effect of agent loss or irregular structures within a swarm. The behaviour causes concave deformities to be removed creating a healing-effect. This technique is implemented by each agent identifying it's position in relation to it's neighbours. If the agent meets a specific criteria the agent's `normal' movement is altered in such a way as to remove a gap between two of it's neighbours. These changes in direction of agents that meet the criterion cause the swarm, as a whole, to be reorganised into a more uniform structure with no internal voids.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.
% Note that keywords are not normally used for peer review papers.
\begin{IEEEkeywords}
Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
\end{IEEEkeywords}
\titlepgskip=-15pt
\maketitle
\section{Introduction}\label{sec:ConcaveReduction}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively.
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two \textit{field effects} to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \textit{Field effects} have ranges around an agent to determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects})the effects are derived from these ranges as vectors. All ranges must fall within the sensing capabilities of the agent ($S_b$). It is usual for the cohesion field to have a radius ($C_b$) which is larger than the repulsion radius ($R_b$). When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will tend towards moving away from the obstacle.
\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/SwarmStableShape1}
{Stable swarm edges\label{concave:SwarmStableShape1}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=5cm]{figures/SwarmStableShape1}
%% \end{center}
%% \caption{Stable swarm edges\label{concave:SwarmStableShape1}}
%% \end{figure}

@@

\begin{equation}\label{eq:Repulsion1}
v_{r}(b) = 
\frac{1}{|rep(b)|}
\left(
\mathlarger{\mathlarger{\sum_{b' \in rep(b)}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}

\begin{equation}\label{eq:FlyToCentre1}
v_{c}(b) = \frac{1}{|nbr(b)|}\left({\mathlarger{\sum_{b' \in nbr(b)}}{bb'}}\right)
\end{equation}

\begin{equation}\label{eq:AgentMovement}
b_{pos}=s_{b}t\big(v(b)\big)\string^ 
\end{equation}

If a swarm is goal-based, the addition of a \emph{directional vector} will prevent all agents simultaneously producing null vectors. Here, $k_dv_{d}(b)$ is the calculated directional vector to be applied to an agent.

\begin{equation}\label{eq:InterAgentMovementDirectional1}
v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
\end{equation}

\begin{equation}
\label{eq:CohesionRepulsion}
P(b) = \left\{\begin{array}{lll}
   |v(b)|& \mathrm{if} & |k_cv_c(b)| > |k_r v_r(b)|\\
  -|v(b)|& \mathrm{otherwise}
  \end{array}\right.
\end{equation}

Obstacles, like agents, can be represented as a point in the system. As an agent moves it may enter an obstacle's \textit{obstacle repulsion field} causing the agent to move away.
In this thesis agents are modelled with a fixed obstacle repulsion distance $O_b$ where a repulsion vector is applied. The repulsion is then a vector of magnitude $O_b$. If more than one obstacle is within the field effect agent the total repulsion vector is the sum of the repulsion vectors due to each obstacle~Figure~\ref{methods:Obstacle1}. The result is normalised and scaled such that the magnitude is the same as the field distance~$O_b$.
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{figures/Obstacle1}
\end{center}
\caption{Obstacle repulsion \label{methods:Obstacle1}}
\end{figure}

Equation~\ref{eq:Obstacle2} shows the resultant repulsion vector $v_o(b)$ for an agent. $\{o~\in~O~:~\magn{bo}~<=~O_b\}$ is the set of obstacles that are within range of agent $b$. $O$ is the set of obstacles. The obstacles are identified using the distance between an agent and an obstacle $\magn{bo}$ and comparing the result to the fixed obstacle repulsion range~$O_b$. The result is calculated by scaling the normalised sum of the normalised vectors $(ob)\string^$ by $O_b$. Note that \string^ is the equivalent of $\hat{v} = \frac{v}{\magn{v}}$ the normalised vector.

\begin{equation}\label{eq:Obstacle2}
v_o(b) =
 O_b\Bigg(\mathlarger{\mathlarger{\mathlarger{\sum}}}_{o \in O~:~\magn{ob} <= O_b}(ob)\string^\Bigg)\string^
\end{equation}

An agent's \textit{movement-destination vector} is the sum of all the component vectors ($v_c, v_r, v_d, v_o$) (Equation~\ref{eq:BotDirection1})~\cite{HAY:08}. For a vector to be used for movement it must have a magnitude of 1 before the agent's speed can be applied.
%%  as shown in~(\autoref{eq:BotDirection1}).

\begin{equation}
\label{eq:BotDirection1}
v(b) = v_{c}(b) + v_{r}(b) + v_{d}(b) + v_{o}(b)
\end{equation}

\begin{equation}\label{eq:BotPhysics1}
v(b) = k_cv_c(b) + k_rv_r(b) + k_dv_d(b) + k_ov_o(b)
\end{equation}

\section{Related Work}\label{sec:RelatedWork}
Research have identified several approaches to resolving swarm structure issues. A prototype framework for self healing swarms was developed by Dai et al. which considered the problem of agent failure in hostile environments~\cite{DHMRZ:06}. This is in line with work carried out by Vassev and Hinchey who model swarm deployment using ASSL (Autonomic System Specification Language)~\cite{VH:09}. This technique was used by NASA (National Aeronautics and Space Administration) when developing their ANTS (Autonomous Nano Technology Swarm) for use in asteroid belt exploration. This line of research is more focused towards an agent's internal systems failure rather than the removal of anomalies in a swarm distribution. Roach et al. takes a different approach and focuses on the effects of sensor failure and the effect that has on agent distribution~\cite{RMT:15}. The closest research to concave reduction, as discussed in this paper, is that of Lee and Chong who identify the issue of concave edges within swarms in an attempt to create regular lattice formations~\cite{GN:08}. The main focus of their paper is the restructuring of the internal distribution of inter-agent formations. The research by Ismail and Timmis used the approach of \textit{bio-inspired} healing using \textit{granuloma formation}, a biological method for encapsulating an antigen~\cite{IT:10}.
Concave reduction is an extension of the work discussed by Ismail and Timmis in their paper on `self-healing' \cite{IT:10} and Lee and Chong's work on identifying concave edges~\cite{GN:08}. The technique also draws on the work of McLurkin and Demaine who have developed algorithms to detect perimeter types~\cite{MD:09}, although this paper does not identify the perimeter types as this would require a communications infrastructure.

\section{Perimeter Detection}\label{sec:PerimeterDetection}
%In a 2D Euclidean plane a circle provides the maximum area with the minimum perimeter size. A circles perimeter is convex in nature. 
In a disorganised swarm agents that are on a perimeter edge need to be identified as the starting point of coordinating the structure. The edges can form part of an outer (green) or inner edge (red) (Fig.~\ref{fig:PerimeterBots1}) or be part of an island~(Fig.~\ref{fig:PerimeterBots2}).
Agents are identified as being perimeter-based if they comply with a set of conditions. These conditions are: 
\begin{itemize}
  \item There are at least two adjacent neighbours that cannot detect each other.
  \item The agent has less than 4 neighbours.
  \item The angle between two neighbours is $> 180^\circ$.
\end{itemize}

These conditions create an identifiable set of agents that create a perimeter.
This process detects both the outer edge of a swarm and any internal features~(voids) that satisfy the same set of conditions. It is therefore possible to have both voids and islands of agents within the same swarm~(Fig.~\ref{fig:PerimeterBots2}). Voids are best defined as perimeters that are concave in nature and inside another perimeter. McLurkin describes two types of perimeters convex and concave~\cite{MD:09}. A convex perimeter is an edge where the average angle of the agent's exposed faces are~$< 180^\circ$. A concave perimeter is where the average exposed angle is~$> 180^\circ$.

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/PerimeterBots2}
{Inner Island Perimeter\label{fig:PerimeterBots2}}
%\subsection{Full-perimeter detection}\label{sec:PerimeterAgentDetection}

The process of detecting perimeter agents involves several steps. Primarily it involves identifying agents that are not surrounded by interconnected neighbours, as shown in Fig.~\ref{reduced:Perimeter1}, where $b^0$ and $b^1$ are outside of each others cohesion range and are neighbours of $b$. These processes are discussed in more detail here. 
\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/Perimeter1}
{Agent with detected neighbours\label{reduced:Perimeter1}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/Perimeter1}
%% \end{center}
%% \caption{Complex multifaceted agents \label{reduced:Perimeter1}}
%% \end{figure}
The first check is a short circuit count of the number of neighbours. If an agent has $\leq 4$ neighbours and the swarm parameters are for a hexagonal swarm configuration then the agent is a perimeter agent. This identification process is shown in~Fig.~\ref{algo:SmallNeighbourCount}.

@@

If the short circuit \textit{neighbour count} returns false then further conditions must be checked. The additional checks require a dictionary of agent/angular values to be generated as shown in Fig.~\ref{eq:ConcaveVoidPhysics1}. The algorithm shown in Fig.~\ref{algo:getNeighbourAngles} shows how the dictionary of agent/angle values are produced.   
% \begin{equation}
% S_b \buildrel \Delta \over = \{(b',\angle\{(b'~b~b^0)) : b' \in nbr(b)\}\}
% \label{eq:ConcaveVoidPhysics1}
% \end{equation}
Equation~\ref{eq:ConcaveVoidPhysics1} generates the dictionary set of all the neighbours along with an angle that each of the agent's neighbour's make with the first detected neighbour.
%% \begin{center}
%% \begin{equation}
%% \label{eq:ConcaveVoidPhysics2}
%% G_b \buildrel \Delta \over = gap(S_b,\angle<180^\circ)
%% \end{equation}
%% \end{center}

@@

Using the angular dictionary from Fig.~\ref{algo:getNeighbourAngles} it is possible to identify if adjacent neighbours can detect each other. If the agent is within all of its neighbours boundary then it is not on the perimeter~(Fig.~\ref{algo:CheckVisibility}). 
As all the agents have the same field effects the `visibility' of two neighbours can be determined. The angle and distance of each neighbour pair allows the cosine rule to be used to calculate the distance the neighbours are apart as shown in~\ref{eq:CosineRule1} and Fig.~\ref{reduced:Triangulate}. The distance is checked against the neighbour range, if the distance is $\leq N_b$ where $N_b$ is the \textit{neighbour field} then it can be assumed that the neighbours have `sight' of each other.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=3cm]{figures/Triangulate}{Neighbour visibility\label{reduced:Triangulate}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=3cm]{figures/Triangulate}
%% \end{center}
%% \caption{Neighbour visibility\label{reduced:Triangulate}}
%% \end{figure}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/Perimeter2}{Convex agents\label{reduced:Perimeter2}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/Perimeter2}
%% \end{center}
%% \caption{Convex multifaceted agents \label{reduced:Perimeter2}}
%% \end{figure}

@@

\section{Concave Reduction Application}\label{sec:ConcaveReductionApplication}
In a static swarm, where there are no \textit{destination vectors}, concave reduction will result in a restructuring motion that creates a more `rounded' swarm~(Fig.~\ref{fig:OuterPerimeter3}). Concave reduction also creates a surround effect as it removes voids from a swarm~(Section~\ref{voids:ObjectSurrounding}). When concave reduction is applied to a mobile swarm the void closing effect can be used to improve reconnaissance coverage as a swarm passes around obstacles~(Section~\ref{concave:mobileSwarm1}).
Although these effects improve the potential applications of swarms there is a negative impact introduced by concave reduction. In some circumstances concave reduction can create an artificial aggregate \textit{destination vector}. 
To implement void reduction full perimeter detection is required. Concave reduction does not require the perimeter type to be identified and therefore no communications infrastructure is required~\cite{MD:09,NIM:09,SOM:12,ZFG:13,JG:13}. The main limitation a communications infrastructure imposes on a swarm is the number of agents a swarm can contain. This is due to the propagation time of agent positions throughout the swarm. 
The identification of candidate agents on the perimeter is incorporated into the perimeter detection algorithm as shown in Fig.~\ref{algo:checkVisibility2}.\\

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/PerimeterBots4}{Link\label{fig:Connector1}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/PerimeterBots4}
%% \end{center}
%% \caption{Link\label{fig:Connector1}}
%% \end{figure}
\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/PerimeterBots5}{Bridge\label{fig:Connector2}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=4cm]{figures/PerimeterBots5}
%% \end{center}
%% \caption{Bridge\label{fig:Connector2}}
%% \end{figure}

\section{Concave reduction agent movement}\label{concave:AgentMovement}
Adding a further characteristic to the motion of a swarm necessitates a revision to the existing agent model~(Equation.~\ref{eq:InterAgentMovement2}). With concave reduction this revision is based upon the identification of the concave perimeter edges as shown in~Fig.~\ref{concave:VoidConcave1} as $(b^{'},b,b^{''})$. 
Equation~\ref{eq:InterAgentMovement2} identifies the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.

@@

Where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:BaselineConcaveReduction}.\\
% \begin{figure}
% \begin{center}
% \includegraphics[width=5cm]{figures/VoidConcave1}
% \end{center}
% \caption{Agent concave motion\label{concave:VoidConcave1}}
% \end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/VoidConcave1}{Agent concave motion\label{concave:VoidConcave1}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/InterAgentEffect1}{Initial Position\label{fig:InterAgentEffect1}}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=5cm]{figures/InterAgentEffect1}
%% \end{center}
%% \caption{Initial Position\label{fig:InterAgentEffect1}}
%% \end{figure}
